---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-",
  fig.height = 8,
  fig.width = 12
)
```

# auto_cogs
[![Travis-CI Build Status](https://travis-ci.org/schloerke/autocogs.svg?branch=master)](https://travis-ci.org/schloerke/autocogs)
[![Coverage Status](https://img.shields.io/codecov/c/github/schloerke/autocogs/master.svg)](https://codecov.io/github/schloerke/autocogs?branch=master)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/autocogs)](https://cran.r-project.org/package=autocogs)


Automatically add cognostics for each plot in a given panel column.  Cognostics are univariate statistics for a subset of data.  Autocogs are particularly useful for list-column `data_frame`s being fed into `trelliscopejs::trelliscope`.

## Installation

You can install autocogs from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("schloerke/autocogs")
```

## Examples

### Gapminder

This is a basic example which shows you how to solve a common problem:

```{r gapminder}
library(gapminder)
library(tidyverse)
library(autocogs)

# Set up data
## nest the data according to the country and continent
gapminder %>%
  group_by(country, continent) %>%
  nest() %>%
  mutate(
    # create a column of plots with a
    # * line
    # * points
    # * and a linear model
    panel = lapply(data, function(dt) {
      ggplot(dt, aes(year, lifeExp)) +
        geom_line() +
        geom_point() +
        geom_smooth(method = "lm")
    })
  ) %>%
  print() ->
gap_data

# Look at the first panel (ggplot2 plot)
gap_data$panel[[1]]

# Add cognostic information given the panel column plots
full_gap_data <- gap_data %>% add_panel_cogs() %>% print()

# display the panel and cognostics in a trelliscopejs widget
## sort all panels by worst to best R^2 (percent explained by linear model)
trelliscopejs::trelliscope(
  full_gap_data, "gapminder life expectancy",
  panel_col = "panel",
  self_contained = TRUE,
  ncol = 8, nrow = 4,
  auto_cog = FALSE,
  state = list(
    sort = list(list(name = "r2", dir = "asc")),
    labels = c("country", "continent", "r2")
  )
)
# (screen shot of trelliscopejs widget)
```
